---
title: "dbGaP DGE"
output: html_notebook
---

Loading required libraries


```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
library(tibble)
library(limma)
library(xCell)
```

-------------------
Not cell-corrected
-------------------

Create design matrix for DGE analysis

```{r}
design_pooled <- model.matrix(~Disorder + sex + cohort + Age, data = final_metadata)

final_metadata_male <- final_metadata %>% filter(sex == "Male")
design_male <- model.matrix(~Disorder + cohort + Age, data = final_metadata_male)

final_metadata_female <- final_metadata %>% filter(sex =="Female")
design_female <- model.matrix(~Disorder + cohort + Age, data = final_metadata_female)
```

Conduct DGE analysis using limma

```{r}
fit_pooled <- lmFit(eset_fin, design_pooled)
fit_pooled <- eBayes(fit_pooled, trend = T)
dbgap_tt_ncc_pooled <- topTable(fit_pooled, coef = "DisorderMDD", number = 45485)
results <- decideTests(fit_pooled)
summary(results)

eset_fin_m <- eset_fin[,final_metadata_male$sample_id_match]
fit_m <- lmFit(eset_fin_m, design_male)
fit_m <- eBayes(fit_m, trend = T)
dbgap_tt_ncc_m <- topTable(fit_m, coef = "DisorderMDD", number = 45485)
results_m <- decideTests(fit_m)
summary(results_m)

eset_fin_f <- eset_fin[,final_metadata_female$sample_id_match]
fit_f <- lmFit(eset_fin_f, design_female)
fit_f <- eBayes(fit_f, trend = T)
dbgap_tt_ncc_f <- topTable(fit_f, coef = "DisorderMDD", number = 45485)
results_f <- decideTests(fit_f)
summary(results_f)
```

-----------------------
Cell-corrected analysis
-----------------------

Calculate cell proportions using xCell

```{r}
eset_selected <- eset_fin@assayData$exprs %>% as.data.frame() %>% rownames_to_column("PROBEID") %>% left_join(., eset_annotation, by = "PROBEID") %>% column_to_rownames("PROBEID")
unq_dbgap <- collapseRows(datET = eset_selected[,1:4931], rowGroup = eset_selected$SYMBOL, rowID = rownames(eset_selected))
cc_mat <- unq_dbgap$datETcollapsed %>% as.data.frame() 

#load cell count information available for a subset of data
blood_data <- read.table("Sample_Attributes.txt", stringsAsFactors = F, header = T, sep = "\t", skip = 10) %>% dplyr::select(dbGaP_Sample_ID:hct) 
blood_data <- blood_data %>% filter(SAMPID %in% final_metadata$SAMPID)
all_present <- blood_data %>% dplyr::select(SAMPID, Wbc:rbc) %>% na_if(., -1) %>% mutate(na_count = rowSums(is.na(.))) %>% filter(na_count ==0) %>% mutate(sample_id_match = paste("X", SAMPID, sep = ""))


cc_counts <- cc_mat %>% dplyr::select(one_of(dput(all_present$sample_id_match)))
cc_metadata <- final_metadata %>% filter(sample_id_match %in% colnames(cc_counts))
```

Check overlap of at least 5000 gene names (recommended) between xCell database and input

```{r}
xcell_genes <- xCell.data$genes %>% as.data.frame() %>% setNames("gene_name")
cc_genes <- rownames(cc_mat) %>% as.data.frame() %>% setNames("gene_name")

venn.diagram(list(dbGaP_cc_genes = cc_genes$gene_name, xCell = xcell_genes$gene_name), filename = "overlap_genes_with_xCell.tiff")
```

xCell - non normalisation required for microarray data

xCell calculation

```{r}
cellchoice_names <- c("B-cells", "CD4+ T-cells", "CD8+ T-cells", "Monocytes", "NK cells", "Basophils", "Eosinophils", "Neutrophils") 

set.seed(12006)
xcellout_pooled <- xCellAnalysis(expr=as.matrix(cc_mat), genes=rownames(cc_mat),rnaseq=F,
                                 cell.types.use=cellchoice_names)

```

### xCell stats
code credit M-E Lynall

#significance
```{r}
set.seed(12008)
tmp <- xCellSignifcanceBetaDist(xcellout_pooled, rnaseq=F)
toselect <- apply(tmp,1,function(x){sum(x<0.1)}>=1362) #30% samples
toselect %>% table 
xcellsmall_pooled <- xcellout_pooled[toselect,]
ypos_pooled <- apply(xcellsmall_pooled, 1, function(x){1.3*max(x)})
ypos_mono <- apply(xcellout_pooled, 1, function(x){1.3*max(x)})
```


```{r}
xcell_pooled <- melt(xcellsmall_pooled, varnames=c("Celltype","sample"), value.name="proportion")
xcell_pooled$cond <- final_metadata$Disorder[match(xcell_pooled$sample,final_metadata$sample_id_match)]

anno_df<- compare_means(proportion ~ cond, group.by="Celltype",data=xcell_pooled, method="wilcox.test", paired=FALSE, p.adjust.method = "BH") %>%
 dplyr::mutate(p.adj = format.pval(p.adj, digits = 1), p = format.pval(p, digits = 1))
anno_df$y.position <- ypos_pooled
order <- as.character(data.frame(anno_df %>% arrange(p))$Celltype)
print(anno_df)
xcell_pooled$Celltype <- factor(as.character(xcell_pooled$Celltype), levels=order, ordered=T)

#Figure 1
ggplot(xcell_pooled, aes(x=cond, y=proportion)) + 
  geom_boxplot(aes(fill=cond),position=position_dodge(), outlier.shape = NA) + theme_bw() + scale_fill_manual(values = c("cadetblue2","cadetblue")) +
  geom_jitter(color="black", width = 0.2, cex = 0.1) +
  facet_wrap(~Celltype, scales = "free_y", ncol = 4) + stat_pvalue_manual(data=anno_df, label="p.signif", vjust=1.2)  + labs(x = "", y = "", fill = "Disorder")# Note that p.signif is unadjusted p values
ggsave("xcell_prop_pooled.pdf" ,width=9, height=4)
```


```{r}
cc_pooled <- xcellsmall_pooled %>% t() %>% as.data.frame() %>% rownames_to_column("sample_id_match")
cc_pooled <- cc_pooled %>% mutate(Lymphocytes = `B-cells` + `CD4+ T-cells` + `CD8+ T-cells` + `NK cells`)

cc_mono <- xcellout_pooled %>% t() %>% as.data.frame() %>% rownames_to_column("sample_id_match")
cc_mono <- cc_mono %>% mutate(Lymphocytes = `B-cells` + `CD4+ T-cells` + `CD8+ T-cells` + `NK cells`)
```

```{r}
cc_met <- final_metadata %>% merge(., cc_mono, by = "sample_id_match")
cc_met_male <- cc_met %>% filter(sex == "Male")
cc_met_fem <- cc_met %>% filter(sex =="Female")
```

Create design matrix for DGE analysis

```{r}
design_pooled_cc <- model.matrix(~Disorder + Sex + cohort + Age + Lymphocytes + Monocytes + Eosinophils + Neutrophils, data = cc_met)

design_female_cc <- model.matrix(~Disorder + cohort + Age + Lymphocytes + Monocytes + Eosinophils + Neutrophils, data = cc_met_fem)

design_male_cc <- model.matrix(~Disorder + cohort + Age + Lymphocytes + Monocytes + Eosinophils + Neutrophils, data = cc_met_male)
```

Conduct DGE analysis using limma

```{r}
fit_pooled_cc <- lmFit(eset_fin, design_pooled_cc)
fit_pooled_cc <- eBayes(fit_pooled_cc, trend = T)
dbgap_tt_cc_pooled <- topTable(fit_pooled_cc, coef = "DisorderMDD", number = 45485)
results <- decideTests(fit_pooled_cc)
summary(results)

eset_fin_m <- eset_fin[,cc_met_male$sample_id_match]
fit_m_cc <- lmFit(eset_fin_m, design_male_cc)
fit_m_cc <- eBayes(fit_m_cc, trend = T)
dbgap_tt_cc_m <- topTable(fit_m_cc, coef = "DisorderMDD", number = 45485)
results_m <- decideTests(fit_m_cc)
summary(results_m)

eset_fin_f <- eset_fin[,cc_met_fem$sample_id_match]
fit_f_cc <- lmFit(eset_fin_f, design_female_cc)
fit_f_cc <- eBayes(fit_f_cc, trend = T)
dbgap_tt_cc_f <- topTable(fit_f_cc, coef = "DisorderMDD", number = 45485)
results_f <- decideTests(fit_f_cc)
summary(results_f)
```

---------------------------------
Data prep for downstream analysis
---------------------------------

```{r}
catgs <- cc_met_fem %>% dplyr::select(sample_id_match, Disorder, cohort, Age, Eosinophils:Lymphocytes, Monocytes) 
catgs[, c(2,3)] <- lapply(catgs[, c(2,3)], factor)
design = model.matrix(~ cohort + Age + Lymphocytes + Monocytes + Eosinophils + Neutrophils, data = catgs)
mm <- model.matrix(~Disorder, data = catgs)
norm_data <-  unq_dbgap$datETcollapsed %>% as.data.frame() %>% dplyr::select(one_of(dput(cc_met_fem$sample_id_match))) %>% as.matrix()
batch_correct_d_f <- removeBatchEffect(norm_data, covariates = design[,-1], design = mm)

catgs <- cc_met_male %>% dplyr::select(sample_id_match, Disorder, cohort, Age, Eosinophils:Lymphocytes, Monocytes) 
catgs[, c(2,3)] <- lapply(catgs[, c(2,3)], factor)
design = model.matrix(~ cohort + Age + Lymphocytes + Monocytes + Eosinophils + Neutrophils, data = catgs)
mm <- model.matrix(~Disorder, data = catgs)
norm_data <-  unq_dbgap$datETcollapsed %>% as.data.frame() %>% dplyr::select(one_of(dput(cc_met_male$sample_id_match))) %>% as.matrix()
batch_correct_d_m <- removeBatchEffect(norm_data, covariates = design[,-1], design = mm)

catgs <- cc_met %>% dplyr::select(sample_id_match, Disorder, cohort, sex, Age, Eosinophils:Lymphocytes, Monocytes) 
catgs[, c(2,3)] <- lapply(catgs[, c(2,3)], factor)
design = model.matrix(~ cohort + sex + Age + Lymphocytes + Monocytes + Eosinophils + Neutrophils, data = catgs)
mm <- model.matrix(~Disorder, data = catgs)
norm_data <-  unq_dbgap$datETcollapsed %>% as.data.frame() %>% dplyr::select(one_of(dput(cc_met$sample_id_match))) %>% as.matrix()
batch_correct_d_pooled <- removeBatchEffect(norm_data, covariates = design[,-1], design = mm)

metd_d <- cc_met

save(metd_d, batch_correct_d_f, batch_correct_d_m, file = "../for_WGCNA/dbgap.Rdata")
```

--------------------
Sensitivity analysis
--------------------

#BMI
Create design matrix

```{r}

metd_sense <- cc_met %>% mutate(sex = if_else(Sex == 2, "Female", "Male")) %>% filter(is.na(BMI) == F)

design_sens <- model.matrix(~Disorder + sex + cohort + Age + BMI + Lymphocytes + Monocytes + Eosinophils + Neutrophils, data = metd_sense)
design_sens_f <- metd_sense %>% filter(sex =="Female") %>% model.matrix(~Disorder + cohort + Age + BMI + Lymphocytes + Monocytes + Eosinophils + Neutrophils, data = .)
design_sens_m <- metd_sense %>% filter(sex == "Male") %>% model.matrix(~Disorder + cohort + Age + BMI + Lymphocytes + Monocytes + Eosinophils + Neutrophils, data = .)


eset_fin_subset <- eset_fin[,metd_sense$sample_id_match]
check <- metd_sense %>% filter(sex =="Male")
eset_fin_subset_m <- eset_fin_m[,check$sample_id_match]
check <- metd_sense %>% filter(sex =="Female")
eset_fin_subset_f <- eset_fin_f[,check$sample_id_match]

```


```{r}
fit_sens <- lmFit(eset_fin_subset, design_sens)
fit_sens <- eBayes(fit_sens, trend = T)

fit_sens_m <- lmFit(eset_fin_subset_m, design_sens_m)
fit_sens_m <- eBayes(fit_sens_m, trend = T)

fit_sens_f <- lmFit(eset_fin_subset_f, design_sens_f)
fit_sens_f <- eBayes(fit_sens_f, trend = T)

dbgap_tt_sens_pooled <- topTable(fit_sens, coef = "DisorderMDD", number = 45485)
dbgap_tt_sens_f <- topTable(fit_sens_f, coef = "DisorderMDD", number = 45485)
dbgap_tt_sens_m <- topTable(fit_sens_m, coef = "DisorderMDD", number = 45485)
```

#Platelets

```{r}
load("platelet_counts.Rdata", verbose = T)
```


Create design matrix
```{r}

metd_sens_p <- cc_met %>% merge(., platelet_counts, by.x = "sample_id_match", by.y = "X")
design <- model.matrix(~Disorder + Sex + cohort + Age + `B-cells` + `CD4+ T-cells` + `CD8+ T-cells` + Monocytes + `NK cells` + Eosinophils + Neutrophils + Platelets, data = metd_sens_p)
design_male <- metd_sens_p %>% filter(sex == "Male") %>% model.matrix(~Disorder + cohort + Age + `B-cells` + `CD4+ T-cells` + `CD8+ T-cells` + Monocytes + `NK cells` + Eosinophils + Neutrophils + Platelets, data = .)
design_fem <- metd_sens_p %>% filter(sex == "Female") %>% model.matrix(~Disorder + cohort + Age + `B-cells` + `CD4+ T-cells` + `CD8+ T-cells` + Monocytes + `NK cells` + Eosinophils + Neutrophils + Platelets, data = .)
```

DE run

```{r}

fit_sens_p <- lmFit(eset_fin, design)
fit_sens_p <- eBayes(fit_sens_p, trend = T)

fit_sens_p_m <- lmFit(eset_fin_m, design_male)
fit_sens_p_m <- eBayes(fit_sens_p_m, trend = T)

fit_sens_p_f <- lmFit(eset_fin_f, design_fem)
fit_sens_p_f <- eBayes(fit_sens_p_f, trend = T)

dbgap_tt_sens_p_pooled <- topTable(fit_sens_p, coef = "DisorderMDD", number = 45485)
dbgap_tt_sens_p_f <- topTable(fit_sens_p_f, coef = "DisorderMDD", number = 45485)
dbgap_tt_sens_p_m <- topTable(fit_sens_p_m, coef = "DisorderMDD", number = 45485)

```

